Confirmation of Wide-Field Signatures in Redshifted 21~cm Power Spectra using Murchison Widefield Array Observations

### 2015-04-09 ms.tex
Make minor modifications throughout
From 7094fca:
============

In ms.tex,

(*) Make minor modifications throughout

Our recent study \citep{thy15} used all-sky foreground and instrument models for the first time in order to simulate actual EoR experiments more accurately than previous studies. Surprisingly, we found that foreground emission outside the primary beam field of view caused the most significant contamination of the cosmological H~{\sc i} power spectrum Fourier modes that are considered most sensitive for detecting the signal in delay spectrum based analyses. This contamination is the result of the interplay between foreground emission, particularly diffuse Galactic emission, and the wide-field properties typical of EoR instruments. Our simulations predicted that delay spectra from the MWA and other experiments should exhibit a characteristic ``pitchfork'' appearance with local maxima near the horizon delay limits, in addition to at the primary lobe of the power pattern.

@@

In \citet{thy15}, we verified the general features of our simulations against MWA observations, but were unable to confirm the actual {\it pitchfork} prediction due to insufficient sensitivity in the small amount of data analyzed. Here, we use longer MWA integrations to confirm with high significance the presence of the key {\it pitchfork} characteristics of wide-field measurements predicted in the preceding study.

@@

The delay power spectrum is defined as:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frequency of the 21~cm radiation of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard cosmology terms. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, and $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$. $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.

@@

Detailed modeling presented in our original study \citep{thy15} has indicated important implications for future instrument design and analysis algorithms:
\begin{enumerate}
\item positioned adjacent to the Fourier modes most sensitive to the cosmological signal, the {\it pitchfork} feature is a leading source of contamination in these modes, and 
\item of the cases considered, a dish-shaped antenna was found to yield the most desirable response from a foreground contamination viewpoint due to a relatively efficient suppression of sensitivity towards the horizon. 
\end{enumerate}

@@

Figure~\ref{fig:delay-spectra} shows the delay spectra obtained from a single snapshot of MWA data (top), averaging LST aligned delay spectra from 12 individual snapshots from MWA observations on different nights (middle), and from modeling (bottom) with no thermal noise fluctuations \citep{thy15}, for reference. In all panels, the {\it foreground wedge} bounded by horizon limits (white dotted lines) is prominent. The bright branch of power at $\tau=0$ corresponds to foreground emission from the main lobe of the power pattern pointed at the zenith.

@@

Figure~\ref{fig:3-baseline-comparison-delay-spectra} shows the amplitudes of averaged delay spectra on three selected baseline vectors oriented northward. Data and noiseless models are shown in black and red respectively. The horizontal dotted black line denotes {\it rms} of thermal fluctuations estimated from data. The vertical dashed line denotes horizon delay limits, and the vertical dot-dashed lines denote delays at which the responses to coarse band edge flagging are expected.

@@

For future work, we plan to extend our analysis to the upcoming instrument, HERA, which is currently under construction. It is a closely packed hexagonal array of fixed 14~m dishes which will observe the sky drifting overhead with redundant antenna spacings. Based on our earlier study, such a dish will have a much desirable Fourier response from a foreground contamination viewpoint. One of our objectives is to forecast the per-baseline foreground contamination as a function of local sidereal time in order to tune the observing strategy to maximize sensitivity to the cosmological signal.

>>>>

Our recent study \citep{thy15} used all-sky foreground and instrument models for the first time in order to simulate actual EoR experiments more accurately than previous studies. Surprisingly, we found that foreground emission outside the primary beam field of view caused the most significant contamination of the cosmological H~{\sc i} power spectrum Fourier modes that are considered most sensitive for detecting the signal in delay spectrum based analyses. This contamination is the result of the interplay between foreground emission, particularly diffuse Galactic emission, and the wide-field properties typical of EoR instruments. Our simulations predicted that delay spectra from the MWA and other experiments should exhibit a characteristic ``pitchfork'' appearance with local maxima near the horizon delay limits, in addition to at the primary lobe region.

@@

In \citet{thy15}, we verified the general features of our simulations against MWA observations, but were unable to confirm the actual {\it pitchfork} prediction due to insufficient sensitivity in the small amount of data analyzed. Here, we use longer MWA integrations to confirm with high significance the presence of key {\it pitchfork} characteristics of wide-field measurements predicted in the preceding study.

@@

The delay power spectrum is defined as \citep{par12a,thy15}:
\begin{align}\label{eqn:delay-power-spectrum}
  P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel) &\equiv |\tilde{V}_b(\tau)|^2\left(\frac{A_\textrm{e}}{\lambda^2\Delta B}\right)\left(\frac{D^2\Delta D}{\Delta B}\right)\left(\frac{\lambda^2}{2k_\textrm{B}}\right)^2,
\end{align}
with
\begin{align}
  \boldsymbol{k}_\perp &\equiv \frac{2\pi(\frac{\boldsymbol{b}}{\lambda})}{D}, \\
  k_\parallel &\equiv \frac{2\pi\tau\,f_{21}H_0\,E(z)}{c(1+z)^2}, 
\end{align}
where, $A_\textrm{e}$ is the effective area of the antenna, $\Delta B$ is the bandwidth, $\lambda$ is the wavelength of the band center, $k_\textrm{B}$ is the Boltzmann constant, $f_{21}$ is the rest frequency of the 21~cm radiation of H~{\sc i}, $z$ is the redshift, $D\equiv D(z)$ is the transverse comoving distance, $\Delta D$ is the comoving depth along the line of sight, and $h$, $H_0$ and $E(z)\equiv [\Omega_\textrm{M}(1+z)^3+\Omega_\textrm{k}(1+z)^2+\Omega_\Lambda]^{1/2}$ are standard cosmology terms. In this paper, we use $\Omega_\textrm{M}=0.27$, $\Omega_\Lambda=0.73$, $\Omega_\textrm{K}=1-\Omega_\textrm{M}-\Omega_\Lambda$, and $H_0=100\,$km$\,$s$^{-1}\,$Mpc$^{-1}$. $P_\textrm{d}(\boldsymbol{k}_\perp,k_\parallel)$ is in units of K$^2$(Mpc/$h$)$^3$.

@@

Detailed modeling presented in our original study \citep{thy15} has important implications for future instrument design and analysis algorithms:
\begin{enumerate}
\item positioned adjacent to the Fourier modes most sensitive to the cosmological signal, the {\it pitchfork} feature is a leading source of contamination in these modes, and 
\item of the cases considered, a dish-shaped antenna was found to yield the most desirable response from a foreground contamination viewpoint due to a relatively efficient suppression of sensitivity away from the primary field of view. 
\end{enumerate}

@@

Figure~\ref{fig:delay-spectra} shows the delay spectra obtained from a single snapshot of MWA data (top), averaging LST aligned delay spectra from 12 individual snapshots from MWA observations on different nights (middle), and from modeling (bottom) with no thermal noise fluctuations \citep{thy15} shown for reference. In all panels, the {\it foreground wedge} bounded by horizon limits (white dotted lines) is prominent. The bright branch of power at $\tau=0$ corresponds to foreground emission from the main lobe of the power pattern pointed at the zenith.

@@

Figure~\ref{fig:3-baseline-comparison-delay-spectra} shows the averaged delay power spectra on three selected baseline vectors oriented northward. Data and noiseless models are shown in black and red respectively. The horizontal dotted black line denotes {\it rms} of thermal fluctuations estimated from data. The vertical dashed line denotes horizon delay limits, and the vertical dot-dashed lines denote delays at which the responses to coarse band edge flagging are expected.

@@

For future work, we plan to extend our analysis to the upcoming instrument, HERA, which is currently under construction. It is a closely packed hexagonal array of fixed 14~m dishes which will observe the sky drifting overhead with redundant antenna spacings. Based on our earlier study, such a dish will have a much desirable Fourier response from a foreground contamination viewpoint. One of our objectives is to forecast the per-baseline foreground contamination as a function of local sidereal time in order to tune the observing strategy and data analysis to maximize sensitivity to the EoR signal.
